Multi-resolution Neural Field Estimation

### 2011-03-04 Multiresolution_framework/ltx/manuscript.tex
new figure2
a bunch of little comments, edits and suggestions for cutting

\section{Introduction}
\IEEEPARstart{T}{he} human brain has a multiresolution architecture, where spatial scales for information processing and transfer range from individual proteins in synapses, to neurons, to networks of neurons. The multiresolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multiresolution data-driven framework for neural field modeling, the multiresolution approximate integro-difference equation (MRAIDE), to address this challenge.

Neural field models describe the mass action of the central nervous system and are a critical link in our understanding of the biophysics of the EEG. They typically describe the macroscopic dynamics of the human brain, but are also descriptive of finer-scale neurodynamics. The ability to create patient-specific neural models will contribute to our knowledge base of poorly understood diseases such as epilepsy, and furthermore, may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory, but first a suitable model is required. Neural mass and neural field models are the ideal candidates for estimation and control algorithms, due to their strong links with the underlying physiology and their parsimony. It is expected that parameters of the neural fields models will be patient specific and will therefore, need to be inferred from data.

The first work describing data-driven mesoscopic neural modeling~\cite{Valdes1999} used a neural mass model~\cite{LopesDaSilva1976,Zetterberg1978} to fit EEG data. This approach was extended to coupled neural masses (using another model~\cite{Jansen1995}) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM)~\cite{David2003}. Following this work, data-driven modeling was extended to continuum field equations that explain the richer dynamics of spatiotemporal neural fields \cite{Galka2008,schiff2008kalman,Daunizeau2009}. Most recently a framework was developed where a finite element model of the neural field (via a global Galerkin projection) was formed, using a basis function decomposition, to transform the PDE neural field equations into a finite dimension system to facilitate efficient state and parameter estimation~\cite{Freestone2011}. The current paper is an extension to this recent study, where we derive a neural field state-space model that accounts for the multi-resolution architecture and spatial dynamics of the human cortex. In this way, a flexible framework is created where both macroscopic and microscopic behaviour of the system can be represented simultaneously.

@@

where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an i.i.d. disturbance, such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})$. A linear form of the activation function was chosen to construct the estimator due to high computational demands when using a nonlinear (sigmoidal) activation function. The observation equation describing the electrophysiological recordings is given by 

@@

where $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$ is the observation kernel and the observation noise is an i.i.d. white noise process and $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$. % $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel , is corrupted by The superscript $\top$ denotes the transpose operator.

@@

The multi-resolution approximation of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(r)$ and a mother wavelet $\psi(r)$. Considering a one-dimensional field, the connectivity kernel is decomposed by,

@@

where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters, respectively. Scaling functions retain the lowest frequency components, whilst the wavelet functions extract successively higher frequency components (by increasing $j$) due their low and band-pass spectral characteristics, respectively. The field is decomposed as

@@

Eqs.\eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions and must be truncated at some level $j$. The field and the connectivity kernel are approximated by
\begin{align}
	w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu\left(r\right)^\top\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=\left[ \boldsymbol\alpha_{j_0}^\top \quad \boldsymbol\beta_{j_0}^\top \quad \boldsymbol\beta_{j_0+1}^\top \cdots \boldsymbol\beta_{j}^\top\right] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=\left[\mathbf{x}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}+1}^\top \cdots \check{\mathbf{x}}_{t,j}^\top\right].
\label{FieldWeights}
\end{align}
The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$ contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$ respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$ respectively, are defined by
\begin{align}
 \boldsymbol\lambda^\top & (r-r')=\left[ \boldsymbol\phi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0+1}^\top(r-r') \right. \nonumber \\
&\left. \cdots \quad \boldsymbol\psi_{j}^\top(r-r')\right] \label{KernelBasisVector} \\
 \boldsymbol\mu^\top & (r)=\left[ \boldsymbol\phi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0+1}^\top(r) \cdots \boldsymbol\psi_{j}^\top(r) \right]. 
\label{FieldBasisVector}
\end{align}
The vectors in Eqs. (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in Eqs. (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

The MRAIDE can be implemented using various classes of scaling and wavelet functions. B-spline functions were chosen due their compact support and the ability to analytically calculate convolution and inner product to form the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how the matrices are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 

@@

Eq.~\ref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ times convolution of indicator function with itself since

@@

where $q_n$ are the coefficients. By exploiting properties Eqs.~\ref{eq:BsplineConvolution} and \ref{eq:BsplineInnerProduct} the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta} can be computed. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$ the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale. It is important to note that B-spline scaling and wavelet functions have the following orthogonality properties \cite{Unser1993}: 

@@

In this paper $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta}.

@@

The operator $\odot$ denotes block-wise matrix multiplication and  $[.]_{i,j}$ denotes $\left(i,j\right)^{th}$ (block) element of the (block) matrix. Each $1\times n_{\theta}$ block of the $n_x \times n_x$ block matrix,  $\mathbf U$ is given by

@@

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey line) and estimated (black line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, RMSE = 2.10. (b) $j=1$, $n_x=33$, RMSE = 1.42. (c) $j=2$, $n_x=65$, RMSE = 0.80. (d) $j=3$, $n_x=131$, RMSE = 0.72. (e) $j=4$, $n_x=263$, RMSE=0.72.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Kernel.pdf}
		\caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
	\label{fig:KernelEstimate}
\end{figure}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.

>>>>

\section{Introduction}
\IEEEPARstart{T}{he} human brain \mike{cortex?} has a multiresolution architecture, where spatial scales for information processing and transfer range from individual proteins in synapses\cut{, to neurons,} to networks of neurons. This multiresolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multiresolution data-driven framework for neural field modeling, the MultiResolution Approximate Integro-Difference Equation (MRAIDE), to address this challenge.

Neural field models describe the mass action of the central nervous system and are a critical link in our understanding of the biophysics of the EEG. These models typically describe the macroscopic dynamics of the human brain, but are also descriptive of finer-scale neurodynamics. The ability to create patient-specific neural models will contribute to our knowledge base of poorly understood diseases such as epilepsy, and \cut{furthermore,} may also enable the development of new treatment strategies. This is particularly \cut{poignant with}\mike{relevant to} the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using control theory,\mike{which in turn requires a suitable dynamic model.}\cut{ but first a suitable model is required.} Neural mass and neural field models are the ideal candidates for estimation and control algorithms, due to their strong links with the underlying physiology and their parsimony. It is expected that parameters of the neural fields models will be patient specific and will therefore, need to be inferred from data.

The first work describing data-driven mesoscopic neural modeling~\cite{Valdes1999} used a neural mass model~\cite{LopesDaSilva1976,Zetterberg1978} to fit EEG data. This approach was extended to coupled neural masses \cut{(using another model~}\cite{Jansen1995}) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM)~\cite{David2003}. Following this work, data-driven modeling was extended to continuum field equations that explain the richer dynamics of spatiotemporal neural fields \cite{Galka2008,schiff2008kalman,Daunizeau2009}. Most recently a framework was developed where a finite element model of the neural field (via a global Galerkin projection) was formed, using a basis function decomposition, to transform the PDE neural field equations into a finite dimension system to facilitate efficient state and parameter estimation~\cite{Freestone2011}. The current paper is an extension to this recent study, where we derive a neural field state-space model that accounts for the multi-resolution architecture and spatial dynamics of the human cortex. In this way, a flexible framework is created where both macroscopic and microscopic behaviour of the system can be represented simultaneously.

@@

where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the \mike{connectivity} kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an i.i.d. disturbance, such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$\mike{:}\cut{, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes} a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r})$. A linear form of the activation function was chosen to construct the estimator due to high computational demands when using a nonlinear (sigmoidal) activation function. The observation equation describing the electrophysiological recordings is given by \mike{should the $r_n$ be $r_{n_y}$ below?} 

@@

where $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$ is the observation kernel and the observation noise is an i.i.d. \mike{Gaussian} white noise process and $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$. % $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel , is corrupted by The superscript $\top$ denotes the transpose operator.

@@

The multi-resolution approximation of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(r)$ and a mother wavelet $\psi(r)$. Considering a one-dimensional field, the connectivity kernel is decomposed using,

@@

where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters, respectively. Scaling functions retain the lowest frequency components, whilst the wavelet functions extract successively higher frequency components (by increasing $j$) due their low and band-pass spectral characteristics, respectively. The field is decomposed using

@@

Eqs.\eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions and must be truncated at some level $j$. The field and the connectivity kernel are approximated by
\begin{align}
	w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
	\label{eq:KernelFiniteExpansion} \\
	v_t\left(r\right) &\approx \boldsymbol\mu\left(r\right)^\top\mathbf{x}_t,
	\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
\boldsymbol\theta^\top &=\left[ \boldsymbol\alpha_{j_0}^\top \quad \boldsymbol\beta_{j_0}^\top \quad \boldsymbol\beta_{j_0+1}^\top \cdots \boldsymbol\beta_{j}^\top\right] 
\label{KernelWeights} \\
\mathbf{x}_{t}^\top &=\left[\mathbf{x}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}+1}^\top \cdots \check{\mathbf{x}}_{t,j}^\top\right].
\label{FieldWeights}
\end{align}
\mike{I think it would be much clearer to simply define these vectors as vectors, rather than hoping that the reader will make the same mapping from the sets defined here to the vectors you actually use.} The kernel and field approximation coefficient vectors, $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$, contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$ contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$ respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$ respectively, are defined by
\begin{align}
 \boldsymbol\lambda^\top & (r-r')=\left[ \boldsymbol\phi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0+1}^\top(r-r') \right. \nonumber \\
&\left. \cdots \quad \boldsymbol\psi_{j}^\top(r-r')\right] \label{KernelBasisVector} \\
 \boldsymbol\mu^\top & (r)=\left[ \boldsymbol\phi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0+1}^\top(r) \cdots \boldsymbol\psi_{j}^\top(r) \right]. 
\label{FieldBasisVector}
\end{align}
\mike{is it worth defining $s=r-r'$ somewhere and using that throughout? It may save some lines!} The vectors in Eqs. (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in Eqs. (\ref{KernelWeights}) and (\ref{FieldWeights}).

To derive the state-space model, we first define the matrices \mike{bold $r$s here and everywhere (or nowhere)?}
\begin{align}\label{eq:Lambdax}
 \mathbf{\Lambda}_{x} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right)\boldsymbol\mu^\top\left(r\right) dr \\
\label{eq:Lambdatheta}
 \mathbf{\Lambda}_{\theta} &\triangleq \int_{\Omega}\boldsymbol\mu\left(r\right) \int_\Omega { 
	   \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right)
	    \boldsymbol\mu^\top\left(r\right)\mathbf{x}_t 
	\, dr'dr}
\end{align}
Eqs.~\eqref{eq:KernelFiniteExpansion} and \eqref{eq:FieldFiniteExpansion} are then substituted into Eq. \eqref{eq:DiscreteTimeModel}, then we \mike{do you mean pre-multiply? - I'm not sure what cross multiply means here} cross-multiply by $\boldsymbol\mu\left(\mathbf r\right)$ and integrate over the space giving
\begin{align}\label{eq:DecomposedModel2} 
	&\mathbf{\Lambda}_{x} \mathbf{x}_{t+1}= 
	\xi\mathbf{\Lambda}_{x} \mathbf{x}_t +T_s \varsigma \mathbf{\Lambda}_{\theta}\mathbf{x}_t +\int_{\Omega}\boldsymbol\mu\left(r\right)e_t\left(r\right)dr.
\end{align}
Cross-multiplying by $\mathbf{\Lambda}_{x}^{-1}$ gives the state transition equation
\begin{align}\label{eq:StateEquation}
 \mathbf x_{t+1} &=\mathbf A(\boldsymbol \theta) \mathbf x_t+ \mathbf w_t\\
\label{eq:A_theta}
 \mathbf A(\boldsymbol \theta) &= (T_s\varsigma\mathbf{\Lambda}_{x}^{-1}\mathbf{\Lambda}_{\theta}+\xi\mathbf I),
\end{align}
where $\mathbf I$ is the identity matrix. The disturbance becomes 
\begin{equation}\label{eq:Disturbance}
\mathbf w_t= \mathbf{\Lambda}_{x}^{-1}\int_{\Omega}\boldsymbol\mu \left(r\right)e_t\left(r\right)dr,
\end{equation}
which is a vector valued, zero-mean normally distributed white noise process with covariance (see \cite{Scerri2009} for proof)
\begin{equation}
\boldsymbol\Sigma_w =\mathbf{\Lambda}_{x}^{-1}\iint\limits_{\boldsymbol\Omega}\boldsymbol\mu\left(r\right) \eta\left(r-r'\right)\boldsymbol\mu^{\top}\left(r'\right)dr'dr\mathbf{\Lambda}_{x}^{-\top}.
\end{equation}
The observation equation of the state-space model is found by substituting decomposition \eqref{eq:FieldFiniteExpansion}
 into equation \eqref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where each element of the observation matrix is given by
\begin{equation}
	\mathbf{C}_{ij} \triangleq \int_{\Omega}m(r_i - r')\boldsymbol{\mu}_j(r') \, d\mathbf{r}'.
\end{equation}

The MRAIDE can be implemented using various classes of scaling and wavelet functions. B-spline functions were chosen due their compact support and the ability to analytically calculate \cut{the convolution and inner product to form} the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how the matrices are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 

@@

Eq.~\ref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ convolutions of the indicator function with itself since

@@

where $q_n$ are the coefficients. By exploiting \cut{properties} Eqs.~\ref{eq:BsplineConvolution} and \ref{eq:BsplineInnerProduct} the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta} can be computed analytically. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$ the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale\mike{\cite{Goswami1999}}.\cut{ It is important to note that B-spline scaling and wavelet functions have the following orthogonality properties \cite{Unser1993}: 

@@

In this paper $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta}.}

@@

The operator $\odot$ denotes block-wise matrix multiplication \mike{this still doesn't work! block matrix multiplication isn't a different thing to regular matrix multiplication I think? Isn't it just a different way to do regular matrix multiplication? Also one still doesn't know if the indices in eqs 31-33 are block indices or element indices. Also it's hard to invent notation, and if you do you need to make sure it's defined! Maybe check out the tensor literature? (or simply just index all this properly)}and  $[.]_{i,j}$ denotes $\left(i,j\right)^{th}$ (block) element of the (block) matrix. Each $1\times n_{\theta}$ block of the $n_x \times n_x$ block matrix,  $\mathbf U$ is given by

@@

The actual connectivity kernel and the decomposition is plotted in \figurename{\ref{fig:KernelEstimate}}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel lies inside the confidence interval. The small error in the estimate is attributed to the MRA of the system used to form the estimator.
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/Field.pdf}
		\caption{Actual (grey line) and estimated (black line) neural fields at a time instant for different spatial resolutions. (a) $j=0$, $n_x=17$, RMSE = 2.10. (b) $j=1$, $n_x=33$, RMSE = 1.42. (c) $j=2$, $n_x=65$, RMSE = 0.80. (d) $j=3$, $n_x=131$, RMSE = 0.72. (e) $j=4$, $n_x=263$, RMSE=0.72.}
	\label{fig:FieldEstimates}
\end{figure} 
\begin{figure}[!h] 
	\centering
		\includegraphics[scale=1]{./Graph/test.pdf}
		\caption{Cconnectivity kernel estimate. (a) The actual kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). (b) Weighted kernel scaling functions at $j=1$. (c) Weighted kernel wavelet functions at $j=1$.}
	\label{fig:KernelEstimate}
\end{figure}
% \caption{The actual connectivity kernel is shown with the black solid line. The estimated kernel and confidence interval is shown by the red line and red shaded region ($\pm2$ std.). The 7 weighted kernel basis functions are shown by the black dashed lines.}
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution FEM representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. This work is significant, as the ability to create patient specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics, such as epilepsy. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling. For example, it is thought that the dynamics and connectivity structure differs at different spatial scales~\cite{Qubbaj2009}.
